---
title: "Food Survey Investigation"
author: "Flemming Wu"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    vertical_layout: scroll
    theme:
      version: 4
      bg: "#101010"
      fg: "#FDF7F7" 
      primary: "#594B5B"
editor_options: 
  chunk_output_type: inline
runtime: shiny
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

```{r load libraries, warning=FALSE, message=FALSE, results = 'hide', eval=TRUE}
load.lib<-c("haven", "sjlabelled", "data.table", "tidyverse", "xml2", "rvest", "shiny", "plotly", "zoo", "chron") #list of packages needed
install.lib <- load.lib[!load.lib %in% installed.packages()] #select those not installed for installation

for (lib in install.lib) { #install packages needed for installation including dependencies
  install.packages(lib, dependencies = TRUE)
}

sapply(load.lib, require, character = TRUE) #load libraries
```

```{r, message=FALSE}
df <- data.table::fread("https://media.githubusercontent.com/media/flemm0/PM566_project/main/data/processed_data.csv")
```

Figures {data-icon="fa-signal"}
===================================== 

### Figure 1

What time and/or day of the week do people generally eat foods high in sugar or saturated fatty acids (fa)?

```{r consumption by time and day of the week}


ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))

server <- function(input, output, ...) {
  output$p <- renderPlotly({
  x = ifelse(input$ingredient == "Sugar", 
             "total_sugars_gm", "total_saturated_fatty_acids_gm")
  
  df$eating_time <- as.POSIXct(df$time_of_eating_occasion_hhmm, format = "%H:%M:%S")

  df %>%
    select(intake_day_of_the_week, intake_day_cat, time_of_eating_occasion_hhmm, eating_time, !! rlang::sym(x)) %>%
    mutate(eating_time = format(round(eating_time, units="hours"), format="%H:%M")) %>%
    group_by(intake_day_cat, eating_time) %>%
    summarize(m = mean(!! rlang::sym(x))) %>%
    arrange(match(intake_day_cat, 
                  c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")),
            eating_time) %>%
    plot_ly(x = ~eating_time, y = ~m, 
            hoverinfo = "text", 
            text = ~paste( 
              paste(intake_day_cat, ":", sep = ""),
              paste("Eating time:", eating_time, "(rounded to nearest hour)", sep= " "),
              paste("Average", tolower(input$ingredient), "consumption (grams):", round(m, digits = 3), sep = " "),
              sep = "<br>"
    )) %>%
    add_lines(color = ~intake_day_cat) %>%
    layout(
      title = paste0("Average ", input$ingredient, " Consumption By Time of Day"),
      yaxis = list(title = "Average Consumption (grams)"),
      xaxis = list(
        title = "Eating Time",
        type = "time",
        tickformat="%H:%M:%S"
        )
      )
  })
}

#zoo::rollmean
shinyApp(ui, server)
```

### Figure 2

```{r testing boxplot, fig.height = 8, fig.width = 12, fig.align = "center"}
ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))

server <- function(input, output, ...){
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    lvls <- df[, eating_occasion := 
              fifelse(eating_occasion == "Desayano", "Desayano (breakfast)",
              fifelse(eating_occasion == "Almuerzo", "Almuerzo (breakfast)",
              fifelse(eating_occasion == "Comida", "Comida (lunch)",
              fifelse(eating_occasion == "Merienda", "Merienda (snack)",
              fifelse(eating_occasion == "Cena", "Cena (dinner)",
              fifelse(eating_occasion == "Entre comida", "Entre comida (snack)",
              fifelse(eating_occasion == "Botana", "Botana (snack)",
              fifelse(eating_occasion == "Bocadillo", "Bocadillo (snack)",
              fifelse(eating_occasion == "Tentempie", "Tentempie (snack)",
              fifelse(eating_occasion == "Bebida", "Bebida (drink)", eating_occasion))))))))))] %>%
  group_by(eating_occasion) %>%
  summarise(m = median(!! rlang::sym(x))) %>%
  arrange(m) %>%
  pull(eating_occasion)
    
  df[, eating_cat := 
     fifelse(grepl("snack", eating_occasion, ignore.case = TRUE), "snack", 
             fifelse(grepl("(breakfast|lunch|dinner|supper|brunch)", eating_occasion, ignore.case = TRUE), 
                     "meal", "other"))] %>%
    mutate(consumption = sqrt(!!rlang::sym(x))) %>%
    select(eating_cat, eating_occasion, consumption) %>%
    plot_ly(x = ~consumption, y = ~factor(eating_occasion, lvls), color = ~eating_cat) %>%
    add_boxplot() %>%
    layout(
      title = paste("Average", input$ingredient, "Consumption By Eating Occasion", sep = " "),
      yaxis = list(title = "Eating Occasion"),
      xaxis = list(title = paste("Square Root", input$ingredient, "Consumption", sep = " "))
    )
  })
}

shinyApp(ui, server)
```

Although eating occasions at different times of the day, such as breakfast, lunch, and dinner do not see much difference in sugar or saturated fatty acid consumption, the plots above do reveal that average sugar consumption is lowest in formal meals such as breakfast, lunch or dinner. On the other end, snacking occasions typically involve much higher average sugar consumption. The reverse trend can be seen for average saturated fa consumption, as snacks are lower on the y-axis than meals such as lunch or dinner.

### Figure 3

Does sugar / saturated fatty acid consumption vary by age, ethnicity, gender?

```{r consumption by age and gender}
ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))

server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    
    df %>%
      mutate(consumption = sqrt(!! rlang::sym(x))) %>%
      select(consumption, 
         age_category, 
         gender, 
         respondent_sequence_number, 
         interview_day) %>%
  plot_ly(x = ~consumption, 
    y = ~factor(interaction(df$age_category, df$gender), 
          levels = c(
                     "<1.male", "<1.female",
                     "1-3.male", "1-3.female",
                     "4-8.male", "4-8.female",
                     "9-13.male", "9-13.female", 
                     "14-18.male", "14-18.female",
                     "19-30.male", "19-30.female",
                     "31-50.male", "31-50.female",
                     "51-70.male", "51-70.female",
                     "70+.male", "70+.female")
          ), 
          color = ~gender, colors = c("#33FFA2", "#FFC229")) %>%
  add_boxplot() %>%
  layout(
    yaxis = list(title = "Age (years)")
    )
  })
}

shinyApp(ui, server)
```

### Figure 4

```{r source of food}
ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))

server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    
    lvls <- df %>%
      group_by(food_source) %>%
      summarise(m = median(!! rlang::sym(x))) %>%
      arrange(m) %>%
      pull(food_source)
    
    df %>%
      filter(!is.na(food_source) & !is.na(!! rlang::sym(x))) %>%
      mutate(consumption = sqrt(!! rlang::sym(x))) %>%
      plot_ly(x = ~consumption, y = ~factor(food_source, lvls)) %>%
      add_boxplot() %>%
      layout(
        yaxis = list(
          title = "Source of Food", 
          tickfont = list(size = 8)
          ),
        xaxis = list(
          title = paste("Square Root of", input$ingredient, "Consumption", sep = " ")
          )
      )
  })
}

shinyApp(ui, server)
```



Supplementary Figures {data-icon="fa-plus"}
===================================== 

### Supplementary Figure 1
Does the source of the food or whether the meal was eaten at home have an effect?

```{r plot meal eaten at home}

ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient",
    label = "Select one:",
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))


server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
               "total_sugars_gm", "total_saturated_fatty_acids_gm")
    df %>%
    filter(!! rlang::sym(x) != 0) %>%
    mutate(consumption = log(!! rlang::sym(x))) %>%
    plot_ly(x = ~consumption, type = "histogram", color = ~did_you_eat_this_meal_at_home)
  })
}

shinyApp(ui, server)
```


### Supplementary Figure 2

What foods are associated with high sugar or high saturated fatty acids?
```{r food items bar chart}

ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient",
    label = "Select one:",
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))



server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    
    if (x == "total_sugars_gm") {
      df %>%
        filter(total_sugars_gm > 33.11) %>%
        count(long_food_code_description) %>%
        top_n(30) %>%
        arrange(desc(n)) %>%
        plot_ly(x = ~n, y = ~long_food_code_description) %>%
        add_bars() %>%
        layout(
          yaxis = list(categoryorder = "total ascending",
                            tickfont = list(size = 7),
                            title = "Food Item"),
          xaxis = list(title = "Number of Times Recorded in Dataset")
          )
    } else {
      df %>%
        filter(total_saturated_fatty_acids_gm > 8.511) %>%
        count(long_food_code_description) %>%
        top_n(30) %>%
        arrange(desc(n)) %>%
        plot_ly(x = ~n, y = ~long_food_code_description) %>%
        add_bars() %>%
        layout(
          yaxis = list(categoryorder = "total ascending",
                            tickfont = list(size = 7),
                            title = "Food Item"),
          xaxis = list(title = "Number of Times Recorded in Dataset")
          )
    }
      
    
  })
}

shinyApp(ui, server)

```


### Supplementary Figure 3

```{r eval = FALSE}
ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))

server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    
    df %>%
      mutate(consumption = !! rlang::sym(x)) %>%
      select(consumption, 
         racehispanic_origin_w_nh_asian, 
         gender, 
         respondent_sequence_number, 
         interview_day) %>%
      group_by(respondent_sequence_number) %>%
      summarize(consumption = mean(consumption)) %>%
      mutate(consumption = sqrt(consumption)) %>%
      plot_ly(x = ~consumption, y = ~racehispanic_origin_w_nh_asian) %>%
      add_boxplot()
  })
}

shinyApp(ui, server)
```


