---
title: "Food Survey Investigation"
author: "Flemming Wu"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    vertical_layout: scroll
    theme:
      version: 4
      bg: "#101010"
      fg: "#FDF7F7" 
      primary: "#594B5B"
editor_options: 
  chunk_output_type: inline
runtime: shiny
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

```{r}
library(data.table)
library(tidyverse)
library(rvest)
library(shiny)
library(plotly)
library(chron)
```


```{r, message=FALSE}
df <- data.table::fread("https://media.githubusercontent.com/media/flemm0/PM566_project/main/data/processed_data.csv")
```

Figures {data-icon="fa-signal"}
===================================== 

### Figure 1

What time and/or day of the week do people generally eat foods high in sugar or saturated fatty acids (fa)?

```{r consumption by time and day of the week}
ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))

server <- function(input, output, ...) {
  output$p <- renderPlotly({
  x = ifelse(input$ingredient == "Sugar", 
             "total_sugars_gm", "total_saturated_fatty_acids_gm")
  
  df$eating_time <- as.POSIXct(df$time_of_eating_occasion_hhmm, format = "%H:%M:%S")

  df %>%
    select(intake_day_of_the_week, intake_day_cat, time_of_eating_occasion_hhmm, eating_time, !! rlang::sym(x)) %>%
    mutate(eating_time = format(round(eating_time, units="hours"), format="%H:%M")) %>%
    group_by(intake_day_cat, eating_time) %>%
    summarize(m = mean(!! rlang::sym(x))) %>%
    arrange(match(intake_day_cat, 
                  c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")),
            eating_time) %>%
    plot_ly(x = ~eating_time, y = ~m, 
            hoverinfo = "text", 
            text = ~paste( 
              paste(intake_day_cat, ":", sep = ""),
              paste("Eating time:", eating_time, "(rounded to nearest hour)", sep= " "),
              paste("Average", tolower(input$ingredient), "consumption (grams):", round(m, digits = 3), sep = " "),
              sep = "<br>"
    )) %>%
    add_lines(color = ~intake_day_cat) %>%
    layout(
      title = paste0("Average ", input$ingredient, " Consumption By Time of Day"),
      yaxis = list(title = "Average Consumption (grams)"),
      xaxis = list(
        title = "Eating Time",
        type = "time",
        tickformat="%H:%M:%S"
        )
      )
  })
}

shinyApp(ui, server)
```

### Figure 2

```{r testing boxplot, fig.height = 8, fig.width = 12, fig.align = "center"}
ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))

server <- function(input, output, ...){
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    lvls <- df[, eating_occasion := 
              fifelse(eating_occasion == "Desayano", "Desayano (breakfast)",
              fifelse(eating_occasion == "Almuerzo", "Almuerzo (breakfast)",
              fifelse(eating_occasion == "Comida", "Comida (lunch)",
              fifelse(eating_occasion == "Merienda", "Merienda (snack)",
              fifelse(eating_occasion == "Cena", "Cena (dinner)",
              fifelse(eating_occasion == "Entre comida", "Entre comida (snack)",
              fifelse(eating_occasion == "Botana", "Botana (snack)",
              fifelse(eating_occasion == "Bocadillo", "Bocadillo (snack)",
              fifelse(eating_occasion == "Tentempie", "Tentempie (snack)",
              fifelse(eating_occasion == "Bebida", "Bebida (drink)", eating_occasion))))))))))] %>%
  group_by(eating_occasion) %>%
  summarise(m = median(!! rlang::sym(x))) %>%
  arrange(m) %>%
  pull(eating_occasion)
    
  df[, eating_cat := 
     fifelse(grepl("snack", eating_occasion, ignore.case = TRUE), "snack", 
             fifelse(grepl("(breakfast|lunch|dinner|supper|brunch)", eating_occasion, ignore.case = TRUE), 
                     "meal", "other"))] %>%
    mutate(consumption = sqrt(!!rlang::sym(x))) %>%
    select(eating_cat, eating_occasion, consumption) %>%
    plot_ly(x = ~consumption, y = ~factor(eating_occasion, lvls), color = ~eating_cat) %>%
    add_boxplot() %>%
    layout(
      title = paste("Distribution of", input$ingredient, "Consumption By Eating Occasion", sep = " "),
      yaxis = list(title = "Eating Occasion"),
      xaxis = list(title = paste("Square Root", input$ingredient, "Consumption", sep = " "), hoverformat = ".2f")
    )
  })
}

shinyApp(ui, server)
```

Although eating occasions at different times of the day, such as breakfast, lunch, and dinner do not see much difference in sugar or saturated fatty acid consumption, the plots above do reveal that average sugar consumption is lowest in formal meals such as breakfast, lunch or dinner. On the other end, snacking occasions typically involve much higher average sugar consumption. The reverse trend can be seen for average saturated fa consumption, as snacks are lower on the y-axis than meals such as lunch or dinner.

### Figure 3

Does sugar / saturated fatty acid consumption vary by age, ethnicity, gender?

```{r consumption by age and gender (change to barplot with means and also grouped by ethnicity)}
ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))

server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    
    df %>%
      select(racehispanic_origin_w_nh_asian,
             age_category, 
             gender, 
             respondent_sequence_number, 
             interview_day,
             !!rlang::sym(x)) %>%
      group_by(racehispanic_origin_w_nh_asian, age_category, gender) %>%
      summarize(total_consumption = sum(!!rlang::sym(x))) %>%
      group_by(age_category, gender) %>%
      mutate(prop = total_consumption / sum(total_consumption)) %>%
      ungroup() %>%
      unite(age_gender, c(age_category, gender), sep = " ") %>%
      plot_ly(
        y = ~prop, 
        x = ~factor(age_gender, levels = c(
                         "<1 male", "<1 female",
                         "1-3 male", "1-3 female",
                         "4-8 male", "4-8 female",
                         "9-13 male", "9-13 female", 
                         "14-18 male", "14-18 female",
                         "19-30 male", "19-30 female",
                         "31-50 male", "31-50 female",
                         "51-70 male", "51-70 female",
                         "70+ male", "70+ female")), 
        color = ~racehispanic_origin_w_nh_asian, 
        type = "bar") %>% 
      layout(barmode = "stack", 
             xaxis = list(title = "Age Category (in years) and Gender"), 
             yaxis = list(title = paste("Proportion of", input$ingredient, "Consumption", sep = " ")),
             title = paste("Proportion of", input$ingredient, "Consumption by Age, Gender, and Ethnicity", sep = " "))
  })
}

shinyApp(ui, server)
```

### Figure 4

```{r source of food}
ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))

server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    
    lvls <- df %>%
      filter(!is.na(food_source) & !is.na(!! rlang::sym(x))) %>%
      group_by(food_source) %>%
      summarise(m = median(sqrt(!! rlang::sym(x)))) %>%
      arrange(m) %>%
      pull(food_source)
    
    df %>%
      filter(!is.na(food_source) & !is.na(!! rlang::sym(x))) %>%
      mutate(consumption = sqrt(!! rlang::sym(x))) %>%
      plot_ly(x = ~consumption, y = ~factor(food_source, lvls)) %>%
      add_boxplot() %>%
      layout(
        yaxis = list(title = "Source of Food", tickfont = list(size = 8)),
        xaxis = list(title = paste("Square Root of", input$ingredient, "Consumption", sep = " ")),
        title = paste("Distribution of", input$ingredient, "Consumption By Food Source", sep = " ")
      )
  })
}

shinyApp(ui, server)
```



Supplementary Figures {data-icon="fa-plus"}
===================================== 

### Supplementary Figure 1
Does the source of the food or whether the meal was eaten at home have an effect?

```{r plot meal eaten at home}

ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient",
    label = "Select one:",
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))


server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
               "total_sugars_gm", "total_saturated_fatty_acids_gm")
    df %>%
      filter(!! rlang::sym(x) != 0) %>%
      mutate(consumption = log(!! rlang::sym(x))) %>%
      plot_ly(x = ~consumption, type = "histogram", color = ~did_you_eat_this_meal_at_home, alpha = 0.6, nbinsx = 200) %>%
      layout(xaxis = list(title = paste("Log", input$ingredient, "Consumption", sep = " ")),
             yaxis = list(title = "Count"),
             barmode = "overlay",
             title = paste(input$ingredient, "Consumption vs Whether Meal Eaten at Home"))
  })
}

shinyApp(ui, server)
```


### Supplementary Figure 2

What foods are associated with high sugar or high saturated fatty acids?
```{r food items bar chart}

ui <- shinyUI(fluidPage(
  selectizeInput(
    inputId = "ingredient",
    label = "Select one:",
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
))



server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    
    if (x == "total_sugars_gm") {
      df %>%
        filter(total_sugars_gm > 33.11) %>%
        count(long_food_code_description) %>%
        top_n(30) %>%
        arrange(desc(n)) %>%
        plot_ly(x = ~n, y = ~long_food_code_description) %>%
        add_bars() %>%
        layout(
          yaxis = list(categoryorder = "total ascending",
                            tickfont = list(size = 7),
                            title = "Food Item"),
          xaxis = list(title = "Number of Times Recorded in Dataset"),
          title = "Top 30 Food Items Containing Over 33.11 Grams Total Sugar"
          )
    } else {
      df %>%
        filter(total_saturated_fatty_acids_gm > 8.511) %>%
        count(long_food_code_description) %>%
        top_n(30) %>%
        arrange(desc(n)) %>%
        plot_ly(x = ~n, y = ~long_food_code_description) %>%
        add_bars() %>%
        layout(
          yaxis = list(categoryorder = "total ascending",
                            tickfont = list(size = 7),
                            title = "Food Item"),
          xaxis = list(title = "Number of Times Recorded in Dataset"),
          title = "Top 30 Food Items Containing Over 8.51 Grams Total Saturated Fatty Acid"
          )
    }
      
    
  })
}

shinyApp(ui, server)

```



