---
title: "Food Survey Investigation"
author: "Flemming Wu"
date: "`r Sys.Date()`"
output: flexdashboard::flex_dashboard
editor_options: 
  chunk_output_type: inline
runtime: shiny
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

```{r load libraries, warning=FALSE, message=FALSE, results = 'hide', eval=TRUE}
load.lib<-c("haven", "sjlabelled", "data.table", "tidyverse", "xml2", "rvest", "shiny", "plotly", "zoo", "chron") #list of packages needed
install.lib <- load.lib[!load.lib %in% installed.packages()] #select those not installed for installation

for (lib in install.lib) { #install packages needed for installation including dependencies
  install.packages(lib, dependencies = TRUE)
}

sapply(load.lib, require, character = TRUE) #load libraries
```

```{r, message=FALSE}
df <- data.table::fread("https://media.githubusercontent.com/media/flemm0/PM566_project/main/data/processed_data.csv")
```

## Figures

##### What time and/or day of the week do people generally eat foods high in sugar or saturated fatty acids (fa)?


```{r consumption by time and day of the week}
ui <- fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
)

server <- function(input, output, ...) {
  output$p <- renderPlotly({
  x = ifelse(input$ingredient == "Sugar", 
             "total_sugars_gm", "total_saturated_fatty_acids_gm")
  
  df$eating_time <- as.POSIXct(df$time_of_eating_occasion_hhmm, format = "%H:%M:%S")

  df %>%
    select(intake_day_of_the_week, intake_day_cat, time_of_eating_occasion_hhmm, eating_time, !! rlang::sym(x)) %>%
    #mutate(eating_time = floor_date(eating_time, unit="30minutes")) %>%
    group_by(intake_day_cat, eating_time) %>%
    summarize(m = mean(!! rlang::sym(x))) %>%
    arrange(match(intake_day_cat, c("Monday", "Tuesday", "Wednesday", 
                                    "Thursday", "Friday", "Saturday", "Sunday")), eating_time) %>%
    plot_ly(x = ~eating_time, y = ~m) %>%
    add_lines(color = ~intake_day_cat) %>%
    layout(
      xaxis = list(
        type = 'time',
        tickformat="%H:%M:%S"
        )
      )
  })
}

#zoo::rollmean
shinyApp(ui, server)
```


```{r consumption by eating occasion}
ui <- fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
)

server <- function(input, output, ...) {
  output$p <- renderPlotly({
   x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
   
   df[, eating_occasion := 
                        fifelse(eating_occasion == "Desayano", "Desayano (breakfast)",
                        fifelse(eating_occasion == "Almuerzo", "Almuerzo (breakfast)",
                        fifelse(eating_occasion == "Comida", "Comida (lunch)",
                        fifelse(eating_occasion == "Merienda", "Merienda (snack)",
                        fifelse(eating_occasion == "Cena", "Cena (dinner)",
                        fifelse(eating_occasion == "Entre comida", "Entre comida (snack)",
                        fifelse(eating_occasion == "Botana", "Botana (snack)",
                        fifelse(eating_occasion == "Bocadillo", "Bocadillo (snack)",
                        fifelse(eating_occasion == "Tentempie", "Tentempie (snack)",
                        fifelse(eating_occasion == "Bebida", "Bebida (drink)", eating_occasion))))))))))] %>%
  filter(!is.na(eating_occasion) & !is.na(!! rlang::sym(x))) %>%
  group_by(eating_occasion) %>%
  summarise(m = mean(!! rlang::sym(x)), sd = sd(!! rlang::sym(x))) %>%
  mutate(eating_occasion = forcats::fct_reorder(eating_occasion, m)) %>%
  plot_ly(x = ~m, y = ~eating_occasion) %>%
  add_markers(
    error_x = list(array = ~sd),
    color = I("black"),
    hoverinfo = "x"
    )
  })
}

shinyApp(ui, server)
```





```{r testing boxplot}
ui <- fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
)

server <- function(input, output, ...){
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    lvls <- df[, eating_occasion := 
              fifelse(eating_occasion == "Desayano", "Desayano (breakfast)",
              fifelse(eating_occasion == "Almuerzo", "Almuerzo (breakfast)",
              fifelse(eating_occasion == "Comida", "Comida (lunch)",
              fifelse(eating_occasion == "Merienda", "Merienda (snack)",
              fifelse(eating_occasion == "Cena", "Cena (dinner)",
              fifelse(eating_occasion == "Entre comida", "Entre comida (snack)",
              fifelse(eating_occasion == "Botana", "Botana (snack)",
              fifelse(eating_occasion == "Bocadillo", "Bocadillo (snack)",
              fifelse(eating_occasion == "Tentempie", "Tentempie (snack)",
              fifelse(eating_occasion == "Bebida", "Bebida (drink)", eating_occasion))))))))))] %>%
  group_by(eating_occasion) %>%
  summarise(m = median(!! rlang::sym(x))) %>%
  arrange(m) %>%
  pull(eating_occasion)
    
  df %>%
    mutate(consumption = sqrt(!!rlang::sym(x))) %>%
    plot_ly(x = ~consumption, y = ~factor(eating_occasion, lvls)) %>%
    add_boxplot()
  })
}

shinyApp(ui, server)
```




Although eating occasions at different times of the day, such as breakfast, lunch, and dinner do not see much difference in sugar or saturated fatty acid consumption, the plots above do reveal that average sugar consumption is lowest in formal meals such as breakfast, lunch or dinner. On the other end, snacking occasions typically involve much higher average sugar consumption. The reverse trend can be seen for average saturated fa consumption, as snacks are lower on the y-axis than meals such as lunch or dinner.

#### Does sugar / saturated fatty acid consumption vary by age, ethnicity, gender?

```{r consumption by age and gender}
ui <- fluidPage(
  selectizeInput(
    inputId = "ingredient", 
    label = "Select one:", 
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
)

server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
              "total_sugars_gm", "total_saturated_fatty_acids_gm")
    
    df %>%
      mutate(consumption = sqrt(!! rlang::sym(x))) %>%
      select(consumption, 
         age_category, 
         gender, 
         respondent_sequence_number, 
         interview_day) %>%
  plot_ly(x = ~consumption, 
    y = ~factor(interaction(df$age_category, df$gender), 
          levels = c(
                     "<1.male", "<1.female",
                     "1-3.male", "1-3.female",
                     "4-8.male", "4-8.female",
                     "9-13.male", "9-13.female", 
                     "14-18.male", "14-18.female",
                     "19-30.male", "19-30.female",
                     "31-50.male", "31-50.female",
                     "51-70.male", "51-70.female",
                     "70+.male", "70+.female")
          ), 
          color = ~gender, colors = c("#33FFA2", "#FFC229")) %>%
  add_boxplot() %>%
  layout(yaxis = list(title = "Age"))
  })
}

shinyApp(ui, server)
```




#### Does the source of the food or whether the meal was eaten at home have an effect?

```{r plot meal eaten at home}

ui <- fluidPage(
  selectizeInput(
    inputId = "ingredient",
    label = "Select one:",
    choices = c("Sugar", "Saturated Fatty Acid"),
    selected = "Sugar"
  ),
  plotlyOutput(outputId = "p")
)


server <- function(input, output, ...) {
  output$p <- renderPlotly({
    x = ifelse(input$ingredient == "Sugar", 
               "total_sugars_gm", "total_saturated_fatty_acids_gm")
    df %>%
    filter(!! rlang::sym(x) != 0) %>%
    mutate(consumption = log(!! rlang::sym(x))) %>%
    plot_ly(x = ~consumption, type = "histogram", color = ~did_you_eat_this_meal_at_home)
  })
}

shinyApp(ui, server)
```


```{r plot source of food, warning=FALSE, fig.height = 8, fig.width = 12, fig.align = "center", eval=FALSE}
#sugars
p1 <- df[!is.na(food_source) & !is.na(total_sugars_gm)] %>%
  ggplot(mapping = aes(x = forcats::fct_reorder(factor(food_source), 
                                                total_sugars_gm, mean), y = total_sugars_gm)) +
  stat_summary(fun.data = mean_se, geom = "errorbar") +
  stat_summary(fun = mean, size = 0.4) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
  labs(x = "Source of Food", y = "Total Sugar Consumption (grams)", title = "Sugar Consumption vs Source of Food")


#saturated fa
p2 <- df[!is.na(food_source) & !is.na(total_sugars_gm)] %>%
  ggplot(mapping = aes(x = forcats::fct_reorder(factor(food_source),
                                                           total_saturated_fatty_acids_gm, mean), y = total_saturated_fatty_acids_gm)) +
  stat_summary(fun.data = mean_se, geom = "errorbar") +
  stat_summary(fun = mean, size = 0.4) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
  labs(x = "Source of Food", y = "Total Saturated FA Consumption (grams)", title = "Saturated FA Consumption vs Source of Food")

ggplotly(p1)
```


#### What foods are associated with high sugar or high saturated fatty acids?


```{r, fig.height = 8, fig.width = 12, fig.align = "center",eval=FALSE}

# sugar
p1 <- df[total_sugars_gm > 33.11, .N, by = "long_food_code_description"][order(-N)] %>%
  head(n = 30) %>%
  ggplot(aes(x = N, y = forcats::fct_reorder(long_food_code_description, N))) +
  geom_col() +
  labs(x = "Number of Occurrences",
       y = "Name of Individual Food or Drink Item",
       title = "Top 30 Food or Drink Items Containing Over 33.11 Grams Total Sugar")


# saturated fa
p2 <- df[total_saturated_fatty_acids_gm > 8.511, .N, by = "long_food_code_description"][order(-N)] %>%
  head(n = 30) %>%
  ggplot(aes(x = N, y = forcats::fct_reorder(long_food_code_description, N))) +
  geom_col() +
  labs(x = "Number of Occurrences",
       y = "Name of Individual Food or Drink Item",
       title = "Top 30 Food or Drink Items Containing Over 8.51 Grams Total Saturated Fatty Acids")


ggplotly(p1)
```


